### 2020-02-11 README.Rmd
Fixes.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# listdown

<!-- badges: start -->
[![Build Status](https://travis-ci.org/kaneplusplus/listdown.svg?branch=master)](https://travis-ci.org/kaneplusplus/listdown)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/kaneplusplus/listdown?branch=master&svg=true)](https://ci.appveyor.com/project/kaneplusplus/listdown)
[![Codecov test coverage](https://codecov.io/gh/kaneplusplus/listdown/branch/master/graph/badge.svg)](https://codecov.io/gh/kaneplusplus/listdown?branch=master)
<!-- badges: end -->

The goal of listdown is to programmatically create R Markdown files from 
named lists. The package currently supports the creation of workflowr, 
pdf, word, html, and slide documents.

## Installation

<!--
You can install the released version of listdown from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("listdown")
```

And the development version from [GitHub](https://github.com/) with:
-->

You can install the development version of listdown from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kaneplusplus/listdown")
```

<!-- 
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(listdown)
## basic example code
```
-->