### 2020-10-17 inst/jss/supplemental-material/trial-report.rmd

---
title: Simple Trial Report
output: html_document
---

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
library(gtsummary)
library(flextable)
library(DT)
library(ggplot2)

cc_list <- readRDS("surv-cc.rds")
```

# Table 1

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
cc_list$`Table 1`
```

# Survival Plots

## Overall {.tabset}

### Plot

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
cc_list$`Survival Plots`$`Overall {.tabset}`$Plot
```

### Data

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Overall {.tabset}`$Data)
```

### Table

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Overall {.tabset}`$Table)
```

## Stage {.tabset}

### Plot

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
cc_list$`Survival Plots`$`Stage {.tabset}`$Plot
```

### Data

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Stage {.tabset}`$Data)
```

### Table

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Stage {.tabset}`$Table)
```

## Grade {.tabset}

### Plot

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
cc_list$`Survival Plots`$`Grade {.tabset}`$Plot
```

### Data

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Grade {.tabset}`$Data)
```

### Table

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 4.5}
datatable(cc_list$`Survival Plots`$`Grade {.tabset}`$Table)
```